---
title: "How's The Weather?"
subtitle: "Building and developing a weather forecast system, and applying mechanistic interpretability"
author: "Jeremy Seebohm"
date: today
format: 
  html:
    theme: journal
    toc: true
    toc-depth: 3
    code-fold: true
    embed-resources: true
---

```{=html}
<style>
.glossary-term {
    color: #c44536;
    text-decoration: underline;
    text-decoration-style: dotted;
    cursor: help;
    position: relative;
    font-weight: 500;
}

.glossary-term:hover {
    color: #8b2e23;
    text-decoration-style: solid;
}

.tooltip {
    visibility: hidden;
    position: absolute;
    z-index: 1000;
    background-color: #c44536;
    color: white;
    padding: 12px 15px;
    border-radius: 6px;
    font-size: 0.9em;
    width: 300px;
    bottom: 125%;
    left: 50%;
    transform: translateX(-50%);
    opacity: 0;
    transition: opacity 0.3s;
    box-shadow: 0 4px 12px rgba(0,0,0,0.3);
    line-height: 1.4;
    font-weight: 400;
    text-align: left;
}

.tooltip::after {
    content: "";
    position: absolute;
    top: 100%;
    left: 50%;
    margin-left: -8px;
    border-width: 8px;
    border-style: solid;
    border-color: #c44536 transparent transparent transparent;
}

.glossary-term:hover .tooltip {
    visibility: visible;
    opacity: 1;
}

.intro-box {
    background: #fef2f2;
    border-left: 4px solid #c44536;
    padding: 20px;
    margin: 20px 0;
}

.glossary-section {
    background: #f9f9f9;
    padding: 30px;
    margin-top: 40px;
    border-radius: 8px;
}

.glossary-entry {
    margin-bottom: 20px;
    padding-bottom: 15px;
    border-bottom: 1px solid #ddd;
}

.glossary-entry:last-child {
    border-bottom: none;
}

.glossary-entry strong {
    color: #c44536;
    font-size: 1.1em;
}
</style>
```

```{r}
#| label: setup
#| include: false
library(glue)

# Function to create glossary terms with hover tooltips
glossary <- function(term, definition) {
  glue::glue(
    '<span class="glossary-term">{term}<span class="tooltip">{definition}</span></span>'
  )
}

# Pre-define common terms for easy reuse
enzyme_def <- "A biological catalyst that accelerates chemical reactions in living organisms without being consumed in the process"
protein_def <- "A large biomolecule composed of amino acid chains that performs diverse functions in organisms, including catalysis, structure, and signaling"
catalyst_def <- "A substance that increases the rate of a chemical reaction without undergoing permanent chemical change itself"
substrate_def <- "The molecule upon which an enzyme acts, binding to the active site to undergo chemical transformation"
activesite_def <- "The specific region of an enzyme where substrate molecules bind and undergo a chemical reaction"
algorithm_def <- "A step-by-step procedure or formula for solving a problem or completing a task"
recursion_def <- "A programming technique where a function calls itself to solve smaller instances of the same problem"
datastructure_def <- "A specialized format for organizing, processing, and storing data in computer memory"
entropy_def <- "A measure of disorder or randomness in a system; in thermodynamics, it quantifies the amount of thermal energy unavailable for work"
momentum_def <- "The product of an object's mass and velocity, representing its quantity of motion"
```

::: intro-box
**How to use this manual:** Hover your mouse over any blue underlined term to see its definition instantly. All terms are also listed in the complete glossary at the bottom.
:::

# Introduction

Welcome to this interactive technical manual created with Quarto. This document demonstrates professional hover-style tooltips for technical terms. Simply hover your mouse over any highlighted term to see an instant definition popup.

This manual integrates seamlessly with R code and analysis, making it perfect for technical reports and educational materials.

# Biochemistry Fundamentals

## Enzyme Function

An `r glossary("enzyme", enzyme_def)` is a specialized type of `r glossary("protein", protein_def)` that acts as a biological `r glossary("catalyst", catalyst_def)`. Each `r glossary("enzyme", enzyme_def)` contains an `r glossary("active site", activesite_def)` where the `r glossary("substrate", substrate_def)` binds.

The `r glossary("enzyme", enzyme_def)` reduces the activation energy required for the reaction, allowing biological processes to occur at physiologically relevant rates.

## Energy Currency

Cells store energy in molecules like `r glossary("ATP", "Adenosine triphosphate - the primary energy currency of cells, storing and transferring energy for cellular processes")`, which powers most cellular work. 'r glossary("ATP", "Adenosine triphosphate - the primary")'

## Example with R Code

Here's a simple calculation related to enzyme kinetics:

```{r}
#| label: enzyme-kinetics
#| fig-cap: "Michaelis-Menten Kinetics"

# Michaelis-Menten parameters
Vmax <- 100  # Maximum reaction velocity
Km <- 10     # Michaelis constant

# Substrate concentrations
substrate <- seq(0, 100, by = 5)

# Calculate reaction velocity
velocity <- (Vmax * substrate) / (Km + substrate)

# Plot
plot(substrate, velocity, type = "l", col = "blue", lwd = 2,
     xlab = "Substrate Concentration", 
     ylab = "Reaction Velocity",
     main = "Michaelis-Menten Kinetics")
abline(h = Vmax, col = "red", lty = 2)
text(50, Vmax - 5, "Vmax", col = "red")
```

The curve shows how `r glossary("enzyme", enzyme_def)` velocity depends on `r glossary("substrate", substrate_def)` concentration.

# Computer Science Principles

## Algorithm Design

An `r glossary("algorithm", algorithm_def)` is a precise sequence of instructions for solving a problem. One powerful technique is `r glossary("recursion", recursion_def)`, where an `r glossary("algorithm", algorithm_def)` calls itself with modified parameters.

Efficient `r glossary("algorithms", algorithm_def)` often rely on optimal `r glossary("data structures", datastructure_def)` to achieve fast execution times.

## Example: Fibonacci with Recursion

```{r}
#| label: fibonacci

# Recursive implementation
fibonacci_recursive <- function(n) {
  if (n <= 1) return(n)
  return(fibonacci_recursive(n-1) + fibonacci_recursive(n-2))
}

# Calculate first 10 Fibonacci numbers
fib_numbers <- sapply(0:10, fibonacci_recursive)
print(fib_numbers)
```

This demonstrates `r glossary("recursion", recursion_def)` in action - the function calls itself to solve the problem.

# Thermodynamics and Mechanics

## Energy and Disorder

The second law of thermodynamics states that `r glossary("entropy", entropy_def)` tends to increase in isolated systems. This principle explains why processes have preferred directions and why perfect efficiency is impossible.

## Motion and Conservation

When an object moves, it possesses `r glossary("momentum", momentum_def)`, which is conserved in closed systems.

```{r}
#| label: momentum-conservation

# Example: Two objects colliding
m1 <- 5  # mass 1 (kg)
v1 <- 10 # velocity 1 (m/s)
m2 <- 3  # mass 2 (kg)
v2 <- -5 # velocity 2 (m/s, negative = opposite direction)

# Total momentum before collision
momentum_before <- m1 * v1 + m2 * v2

cat("Total momentum before collision:", momentum_before, "kg·m/s\n")
```

The conservation of `r glossary("momentum", momentum_def)` is fundamental to understanding collisions.

# How to Add New Terms

Adding your own glossary terms is simple! Follow these three steps:

## Step 1: Define in Setup Chunk

Add your definition variable to the setup chunk at the top of the document (after the other definitions):

```{r}
#| eval: false
#| echo: true

photosynthesis_def <- "The process by which plants convert light energy into chemical energy"
```

## Step 2: Use in Text

Then use the `glossary()` function in your text. Here's a real working example:

Plants use `r glossary("photosynthesis", "The process by which plants convert light energy into chemical energy")` to produce glucose and oxygen from sunlight, water, and carbon dioxide.

See how the term appears with a hover tooltip? You can either: - Define the term inline: `glossary("term", "definition here")` - Use a pre-defined variable: `glossary("term", term_def)`

## Step 3: Add to Complete Glossary

Add a corresponding entry in the "Complete Glossary" section at the bottom so readers can find all terms in one place.

# Complete Glossary {#glossary}

:::::::::::::: glossary-section
### All Technical Terms

::: glossary-entry
**Active Site**\
The specific region of an enzyme where substrate molecules bind and undergo a chemical reaction.
:::

::: glossary-entry
**Algorithm**\
A step-by-step procedure or formula for solving a problem or completing a task.
:::

::: glossary-entry
**ATP (Adenosine Triphosphate)**\
The primary energy currency of cells, storing and transferring energy for cellular processes.
:::

::: glossary-entry
**Catalyst**\
A substance that increases the rate of a chemical reaction without undergoing permanent chemical change itself.
:::

::: glossary-entry
**Data Structure**\
A specialized format for organizing, processing, and storing data in computer memory.
:::

::: glossary-entry
**Entropy**\
A measure of disorder or randomness in a system; in thermodynamics, it quantifies the amount of thermal energy unavailable for work.
:::

::: glossary-entry
**Enzyme**\
A biological catalyst that accelerates chemical reactions in living organisms without being consumed in the process.
:::

::: glossary-entry
**Momentum**\
The product of an object's mass and velocity, representing its quantity of motion.
:::

::: glossary-entry
**Protein**\
A large biomolecule composed of amino acid chains that performs diverse functions in organisms, including catalysis, structure, and signaling.
:::

::: glossary-entry
**Recursion**\
A programming technique where a function calls itself to solve smaller instances of the same problem.
:::

::: glossary-entry
**Substrate**\
The molecule upon which an enzyme acts, binding to the active site to undergo chemical transformation.
:::
::::::::::::::

# Session Info

```{r}
#| label: session-info
#| echo: false
sessionInfo()
```
